---
title: "Package functions to filter HTSeq counts data"
author: "Jason Serviss"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

An example of how counts are typically filtered using the package functions.

```{r}
library(sp.scRNAseqData)
library(printr)
```

Load the counts data.
This would usually be done with the read.table function but here we use a small test dataset that represents typical input of HTSeq generated counts.

```{r}
sp.scRNAseqData::testingCounts
```

Move the gene names to rownames.
```{r}
testingCounts <- moveGenesToRownames(testingCounts)
testingCounts
```

Remove the .htseq suffix from colnames.
```{r}
testingCounts <- removeHTSEQsuffix(testingCounts)
testingCounts
```

Label samples as singlets or multiplets.
```{r}
testingCounts <- labelSingletsAndMultiplets(testingCounts, LETTERS[1:5])
testingCounts
```

Detect non-genes reported by HTSeq.
```{r}
nonGenes <- detectNonGenes(testingCounts)
nonGenes
testingCounts <- testingCounts[!nonGenes, ]
testingCounts
```

Detect ERCC reads.
```{r}
ercc <- detectERCCreads(testingCounts)
ercc
testingERCC <- testingCounts[ercc, ]
testingCounts <- testingCounts[!ercc, ]
testingERCC
testingCounts
```

Detect low quality genes.
```{r}
lqg <- detectLowQualityGenes(testingCounts, 18)
lqg
testingCounts <- testingCounts[lqg, ]
testingCounts
```

Detect low quality cells (samples).
Low quality cell detection can also be done with `detectLowQualityCells.ERCCfrac`
which uses ERCC reads to detect cells with a high fraction ERCC.
```{r}
lqc.totalCounts <- detectLowQualityCells.totalCounts(testingCounts, mincount = 25)
lqc.housekeeping <- detectLowQualityCells.housekeeping(testingCounts, geneName = "ACTB", quantileCut = 0.01)
lqc <- lqc.totalCounts & lqc.housekeeping
testingCounts <- testingCounts[, lqc]
testingCounts
```

Convert counts to a matrix.
```{r}
testingCounts <- convertCountsToMatrix(testingCounts)
testingCounts
```

